The \Prob

### 2013-03-05 bikes/bikes.tex
bikes: add a graph

Figure~\ref{fig:sol} shows the solution for 2 people sharing 1 bike.
The frames show progress after $0$, $1$, $1.5$, $2$, and $3$ units of time have elapsed, assuming $w=0.5$ and $b=1$ (i.e., biking is twice as fast as walking).
Note that the bike is idle between times $1$ and $2$ -- for a full third of the transit time! -- yet the stick figures arrive in $3$ time units, a $25\%$ improvement over walking speed, and only $50\%$ slower than both participants having a bike of their own.

\newcommand\PA{\ensuremath{\mathsf{i}}}
\newcommand\PB{\ensuremath{\mathsf{ii}}}
\newcommand\PC{\ensuremath{\mathsf{iii}}}
\newcommand\PD{\ensuremath{\mathsf{iv}}}
\newcommand\PE{\ensuremath{\mathsf{v}}}
Formally we say that the {\em bicycle allocation} for this case is that person \PA~(the blue one, if you're reading this electronically or printed in color) rides bike $0$ for distances between $0$ and $1$, and person \PB~(the red one, if same) rides it for distances between $1$ and $2$. For short, we write:
\begin{eqnarray*}
	p_\PA &=& b_0[0,1] \\
	p_\PB &=& b_0[1,2]
\end{eqnarray*}

@@

Without loss of generality, let us say that $d$ is $n$ units of distance away (i.e., the same as the number of people).
Before presenting a general allocation policy, some observations about the problem:
\begin{enumerate}
	\setlength{\itemsep}{-0.5em}
	\item A total of $n^2$ distances will be traveled. A total of $nk$ distances will be biked. It is pointless for bikes to go backwards.
	\item By symmetry, in an optimal solution, each person will bike $k$ distance and walk $n-k$ distance. This also means that fractional distances (less than $1/n$ of the total distance) need not be considered as places to mount/dismount.
	\item No more than $k$ people can ride for any given distance unit $[i,i+1]$. If any fewer than $k$ people rode this way, the bikes would be under-utilised. This allows us to check both validity and optimality.
	\item The optimal (minimum) total travel time is $k/b + (n-k)/w$. The optimal (maximum) average speed is $nbw/(kw+bn-bk)$.
	% TODO: Discuss epsilon.
\end{enumerate}

@@

\newcommand\BA{\cellcolor{WildStrawberry}~\textcolor{White}{$b_0$}}
\newcommand\BB{\cellcolor{NavyBlue}      ~\textcolor{White}{$b_1$}}
\newcommand\BC{\cellcolor{OliveGreen}   ~\textcolor{White}{$b_2$}}
\begin{figure*}[t]
	\begin{subfigure}[b]{0.28\textwidth}
		\begin{tabular}{rp{1.5em}|p{1.5em}|}
		& \multicolumn{2}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} \\
		\hline
		\multicolumn{1}{r|}{\PA} & \BA & \\
		\cline{2-3}
		\multicolumn{1}{r|}{\PB} & & \BA \\
		\cline{2-3}
		\multicolumn{3}{c}{}\\
		\multicolumn{3}{c}{}\\
		\multicolumn{3}{c}{}\\
		\end{tabular}
		\caption{$2$ people, $1$ bike.}
	\end{subfigure}
	\begin{subfigure}[b]{0.33\textwidth}
		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|}
		& \multicolumn{3}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} \\
		\hline
		\multicolumn{1}{r|}{\PA} & \BA & \BA & \\
		\cline{2-4}
		\multicolumn{1}{r|}{\PB} & & \BB & \BB \\
		\cline{2-4}
		\multicolumn{1}{r|}{\PC} & \BB & & \BA \\
		\cline{2-4}
		\multicolumn{4}{c}{}\\
		\multicolumn{4}{c}{}\\
		\end{tabular}
		\caption{$3$ people, $2$ bikes.}
	\end{subfigure}
	\begin{subfigure}[b]{0.38\textwidth}
		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
		& \multicolumn{5}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} \\
		\hline
		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&   & \\
		\cline{2-6}
		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB& \\
		\cline{2-6}
		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC\\
		\cline{2-6}
		\multicolumn{1}{r|}{\PD} &\BB&   &   &\BA&\BA\\
		\cline{2-6}
		\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BB\\
		\cline{2-6}
		\end{tabular}
		\caption{$5$ people, $3$ bikes.}
	\end{subfigure}
	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.}
	\label{fig:grids}
\end{figure*}

The pattern is now evident: Person $i$ picks up a bike at distance $i$ and drops it off at distance $i+k~\mathsf{mod}~n$.
%TODO
%In the notation from the previous section:
%\begin{eqnarray*}
%	p_i  &=&
%\end{eqnarray*}

@@

\begin{theorem}[Bicycle Allocation]
The \prob~is solved optimally when $p_i$ rides a bike between distance $i$ and distance $i+k~\mathsf{mod}~n$.
\end{theorem}
\begin{proof}
By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride.
% TODO
\end{proof}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

>>>>

\initial{F}igure~\ref{fig:sol} shows the solution for 2 people sharing 1 bike.
The frames show progress after $0$, $1$, $1.5$, $2$, and $3$ units of time have elapsed, assuming $w=0.5$ and $b=1$ (i.e., biking is twice as fast as walking).
Note that the bike is idle between times $1$ and $2$ -- for a full third of the transit time! -- yet the stick figures arrive in $3$ time units, a $25\%$ improvement over walking speed, and only $50\%$ slower than both participants having a bike of their own.

\newcommand\PA{\ensuremath{\mathsf{i}}}
\newcommand\PB{\ensuremath{\mathsf{ii}}}
\newcommand\PC{\ensuremath{\mathsf{iii}}}
\newcommand\PD{\ensuremath{\mathsf{iv}}}
\newcommand\PE{\ensuremath{\mathsf{v}}}
Formally we say that the {\em bicycle allocation} for this case is that person \PA~(the blue one, if you're reading this electronically or printed in color) rides bike $0$ for distances between $0$ and $1$, and person \PB~(the red one, if same) rides it for distances between $1$ and $2$. For short, we write:
\begin{eqnarray*}
	p_\PA &=& b_0@0 \\
	p_\PB &=& b_0@1
\end{eqnarray*}

@@

\initial{W}ithout loss of generality, let us say that $d$ is $n$ units of distance away (i.e., the same as the number of people).
First, some observations about the problem:
\begin{enumerate}
	\setlength{\itemsep}{-0.5em}
	\item A total of $n^2$ distances will be traveled. A total of $nk$ distances will be biked. It is pointless for bikes to go backwards.
	\item By symmetry, in an optimal solution, each person will bike $k$ distance and walk $n-k$ distance. This also means that fractional distances (less than $1/n$ of the total distance) need not be considered as places to mount/dismount.
	\item No more than $k$ people can ride for any given distance unit $[i,i+1]$. If any fewer than $k$ people rode this way, the bikes would be under-utilised. This allows us to check both validity and optimality.
	\item The optimal (minimum) total travel time is $k/b + (n-k)/w$. The optimal (maximum) average speed is $nbw/(kw+bn-bk)$.
	% TODO: Discuss epsilon.
\end{enumerate}

@@

\newcommand\BA{\cellcolor{WildStrawberry}~\textcolor{White}{$b_0$}}
\newcommand\BB{\cellcolor{NavyBlue}      ~\textcolor{White}{$b_1$}}
\newcommand\BC{\cellcolor{OliveGreen}   ~\textcolor{White}{$b_2$}}
\begin{figure*}[t]
	\begin{subfigure}[b]{0.28\textwidth}
		\begin{tabular}{rp{1.5em}|p{1.5em}|}
		& \multicolumn{2}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} \\
		\hline
		\multicolumn{1}{r|}{\PA} & \BA & \\
		\cline{2-3}
		\multicolumn{1}{r|}{\PB} & & \BA \\
		\cline{2-3}
		\multicolumn{3}{c}{}\\
		\multicolumn{3}{c}{}\\
		\multicolumn{3}{c}{}\\
		\end{tabular}
		\caption{$2$ people, $1$ bike.}
	\end{subfigure}
	\begin{subfigure}[b]{0.33\textwidth}
		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|}
		& \multicolumn{3}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} \\
		\hline
		\multicolumn{1}{r|}{\PA} & \BA & \BB & \\
		\cline{2-4}
		\multicolumn{1}{r|}{\PB} & & \BA & \BB \\
		\cline{2-4}
		\multicolumn{1}{r|}{\PC} & \BB & & \BA \\
		\cline{2-4}
		\multicolumn{4}{c}{}\\
		\multicolumn{4}{c}{}\\
		\end{tabular}
		\caption{$3$ people, $2$ bikes.}
	\end{subfigure}
	\begin{subfigure}[b]{0.38\textwidth}
		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
		& \multicolumn{5}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} \\
		\hline
		\multicolumn{1}{r|}{\PA} &\BA&\BB&\BC&   & \\
		\cline{2-6}
		\multicolumn{1}{r|}{\PB} &   &\BA&\BB&\BC& \\
		\cline{2-6}
		\multicolumn{1}{r|}{\PC} &   &   &\BA&\BB&\BC\\
		\cline{2-6}
		\multicolumn{1}{r|}{\PD} &\BC&   &   &\BA&\BB\\
		\cline{2-6}
		\multicolumn{1}{r|}{\PE} &\BB&\BC&   &   &\BA\\
		\cline{2-6}
		\end{tabular}
		\caption{$5$ people, $3$ bikes.}
	\end{subfigure}
	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.
		Can you spot the French flag?}
	\label{fig:grids}
\end{figure*}

The pattern is now evident: Person $i$ rides a bike between distance $i$ and distance $i+k~\mathsf{mod}~n$.
To express this in the notation from the previous section, we can say: $\forall i,j$ with $0\le i < n$ and $0 \le j < k$:
\begin{eqnarray*}
	p_i  &=& b_j @ (i+j~\mathsf{mod}~n)
\end{eqnarray*}

@@

\begin{theorem}[Optimality]
The \prob~is solved optimally when $p_i$ rides between distance $i$ and distance $i+k~\mathsf{mod}~n$.
\end{theorem}
\begin{proof}
By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride.
% TODO
\end{proof}

\subsection{Minimizing Mounts/Dismounts}

\initial{T}he astute reader will note that, while the allocations in the above tables successfully minimize total travel time, they {\em maximize} the number of times each rider must mount and dismount a bicycle. If we let $\epsilon$ (the time to mount or dismount) be non-negligible, we see that the travel time achieved above is actually: $k/b+(n-k)/w + 2k\epsilon$.

We might instead attempt to allocate bicycles contiguously, like so:
		\begin{center}
		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
		& \multicolumn{5}{c}{distance} \\
		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} \\
		\hline
		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&   & \\
		\cline{2-6}
		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB& \\
		\cline{2-6}
		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC\\
		\cline{2-6}
		\multicolumn{1}{r|}{\PD} &\BB&   &   &\BA&\BA\\
		\cline{2-6}
		\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BB\\
		\cline{2-6}
		\end{tabular}
		\end{center}

Here, $n-k+1$ riders can ride ``monogamously'', though $k-1$ people will have two riding  sessions with a walking intermission. This gives us a better travel time of $k/b+(n-k)/w + 4\epsilon$, which is a strict improvement whenever $k>2$.

\begin{theorem}[Contiguous Allocation]
woo
\end{theorem}
\begin{proof}
woo
\end{proof}

\section{Evaluation}

\initial{I}n Figure~\ref{fig:graph} we visualize our findings.
The horizontal line on top represents the state of the art, in which some participants walk the entire distance.
The bottom line represents the ideal case, in which there are enough bikes for everyone to ride together.

Here we have plotted the travel time for various values of $n,k$ against different configurations of the $b/w$ ratio; i.e., how much faster it is to bike than to walk.
We hope this presentation will be useful in many different terrains; for example, if the destination is downhill, $b/w$ may be very high indeed.

\begin{figure}[h]
	\includegraphics[width=0.5\textwidth]{graph.pdf}
	\caption{Speedup achieved in the \prob. As always, lower is better.}
	\label{fig:graph}
\end{figure}
%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
